Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-02-09 Manuscript/Tmatrix.tex
comp details and appendix

In order to target neutral (singly-)excited states, we first consider the static version of the BSE employing the $GT$ quasiparticle energies [see Eq.~\eqref{eq:G0T0}] as well as $T$-matrix kernel [\ie, $\Sigma^{\GT}$ in Eq.~\eqref{eq:Xi}].
In this case, the BSE@$GT$ linear eigenvalue problem simply reads
\begin{equation}
\label{eq:BSE}
	\begin{pmatrix}
		\bA{}{\BSE}		&	\bB{}{\BSE}	\\
		-\bB{}{\BSE}	&	-\bA{}{\BSE}	\\
	\end{pmatrix}
	\cdot
	\begin{pmatrix}
		\bX{m}{\BSE}	\\
		\bY{m}{\BSE}	\\
	\end{pmatrix}
	=
	\Om{m}{\BSE}
	\begin{pmatrix}
		\bX{m}{\BSE}	\\
		\bY{m}{\BSE}	\\
	\end{pmatrix}
\end{equation}
with 
\begin{subequations}
\begin{align}
	A_{ia,jb}^{\BSE} & = \delta_{ij} \delta_{ab} (\e{a}{\GT} - \e{i}{\GT}) + \mel{ib}{}{aj} - T_{ib,aj}(\omega=0)
	\\ 
	B_{ia,jb}^{\BSE} & = \mel{ij}{}{ab} - T_{ij,ab}(\omega=0)
\end{align}
\end{subequations}
The eigenvalues $\Om{m}{\BSE}$ of Eq.~\eqref{eq:BSE} provide $OV$ singlet (\ie, spin-conserved) and $OV$ triplet  (\ie, spin-flip) single excitations.
Note that the spin structure of the BSE@$GT$ equations is analogous to the BSE@$GW$ version, \cite{Monino_2021} and one can compute separately singlet and triplet excitation energies.

Due to the frequency-independent nature of the static BSE, it is well known that one cannot access double (and higher) excitations.
\cite{Loos_2019,Romaniello_2009b,Sangalli_2011,Loos_2020h,Authier_2020,Monino_2021}
In order to go beyond the static approximation, it is possible to consider,  within the dynamical Tamm-Dancoff approximation (which neglects the frequency dependence of the coupling block $\bB{}{}$), the dynamical version of the BSE (dBSE). \cite{Strinati_1988,Romaniello_2009b,Loos_2020h}
In this case, one must solve the (non-linear) dynamical eigenvalue problem
\begin{equation}
\label{eq:dBSE}
	\begin{pmatrix}
		\bA{}{\dBSE}(\Om{S}{})	&	\bB{}{\BSE}	\\
		-\bB{}{\BSE}	&	-\bA{}{\dBSE}(-\Om{S}{})	\\
	\end{pmatrix}
	\cdot
	\begin{pmatrix}
		\bX{S}{\dBSE}	\\
		\bY{S}{\dBSE}	\\
	\end{pmatrix}
	=
	\Om{S}{}
	\begin{pmatrix}
		\bX{S}{\dBSE}	\\
		\bY{S}{\dBSE}	\\
	\end{pmatrix}
\end{equation}
with 
\begin{equation}
	A_{ia,jb}^{\dBSE}(\omega) = \delta_{ij} \delta_{ab} (\e{a}{\GT} - \e{i}{\GT}) + \mel{ib}{}{aj} - \widetilde{T}_{ib,aj}(\omega)
\end{equation}
where, following Strinati's seminal work, \cite{Strinati_1988} one can derive the following expression for the elements of the dynamical $T$ matrix
\begin{equation}
\label{eq:dT}
\begin{split}
	\widetilde{T}_{pq,rs}(\omega) 
		& = \sum_{n} \frac{\braket*{pq}{\chi_n^{N+2}}\braket*{rs}{\chi_n^{N+2}}}{\omega - \Om{n}{N+2} + (\e{i}{\GT} + \e{j}{\GT}) + i\eta}
		\\
		& + \sum_{n} \frac{\braket*{pq}{\chi_n^{N-2}}\braket*{rs}{\chi_n^{N-2}}}{\omega + \Om{n}{N-2} - (\e{a}{\GT} + \e{b}{\GT}) + i\eta}
\end{split}
\end{equation}
from which, one can check that we recover the static expression \eqref{eq:T} in the limit $\Om{n}{N\pm2} \to \infty$.
Equation \eqref{eq:dT} highlights the interesting dynamical structure of the $T$ matrix, where, similarly to the dBSE@$GW$ scheme, \cite{Strinati_1988,Romaniello_2009b,Loos_2020h} the 2h2p configurations are downfolded on the 1h1p configurations. \cite{Bintrim_2021b}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%
The present formalism has been implemented in the electronic structure package \texttt{QuAcK} \cite{QuAcK} that is freely available on \texttt{github}. 
We consider here only systems with closed-shell singlet ground states. 
Thus, the $GW$ and $GT$ calculations are performed by considering a (restricted) HF starting point and standard gaussian basis sets (defined with cartesian functions) are employed.
Note that all quasiparticle energies which are obtained via Eq.~\eqref{eq:G0T0} are corrected in the same way.
Finally, the infinitesimal $\eta$ is set to zero for all calculations.
The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here.
For the sake of completeness, implementable expressions in the spatial orbital basis are reported in the supporting information.

In terms of computational, the overall scaling of BSE@$GT$ is equivalent to BSE@$GW$ as they both correspond to seeking the lowest eigenvalues of a matrix of size $(\Nocc \Nvir \times \Nocc \Nvir)$. 
However, the prefactor corresponding to the calculations of the $T$-matrix and more specifically to the eigenvalue of the \titou{T2: Discuss the computational cost of the BSE@$GT$ equations, especially compare to BSE@$GW$.}
%In terms of computational cost, if one decides to compute the dynamical correction of the $M$ lowest excitation energies, one must perform, first, a conventional (static) BSE calculation and extract the $M$ lowest eigenvalues and their corresponding eigenvectors [see Eq.~\eqref{eq:LR-BSE-stat}].
%These are then used to compute the first-order correction from Eq.~\eqref{eq:Om1-TDA}, which also require to construct and evaluate the dynamical part of the BSE Hamiltonian for each excitation one wants to dynamically correct. 
%The static BSE Hamiltonian is computed once during the static BSE calculation and does not dependent on the targeted excitation.
%
%Searching iteratively for the lowest eigenstates, via Davidson's algorithm for instance, can be performed in $\order*{\Norb^4}$ computational cost.
%Constructing the static and dynamic BSE Hamiltonians is much more expensive as it requires the complete diagonalization of the $(\Nocc \Nvir \times \Nocc \Nvir)$ RPA linear response matrix [see Eq.~\eqref{eq:LR-RPA}], which corresponds to a $\order*{\Nocc^3 \Nvir^3} = \order*{\Norb^6}$ computational cost. 
%Although it might be reduced to $\order*{\Norb^4}$ operations with standard resolution-of-the-identity techniques, \cite{Duchemin_2019,Duchemin_2020} this step is the computational bottleneck in the current implementation.

%For comparison purposes, we employ the theoretical best estimates (TBEs) and geometries of Refs.~\onlinecite{Loos_2018a,Loos_2019,Loos_2020b} from which CIS(D), \cite{Head-Gordon_1994,Head-Gordon_1995} ADC(2), \cite{Trofimov_1997,Dreuw_2015} CC2, \cite{Christiansen_1995a} CCSD, \cite{Purvis_1982} and CC3 \cite{Christiansen_1995b} excitation energies are also extracted. 
%Various statistical quantities are reported in the following: the mean signed error (MSE), mean absolute error (MAE), root-mean-square error (RMSE), and the maximum positive [Max($+$)] and maximum negative [Max($-$)] errors.

@@

@@

%%% FIGURE 2 %%%
\begin{figure*}
	\includegraphics[width=\textwidth]{BeH2}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{BeH2} as a function of the distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels.
	The dynamically-corrected BSE excitation energies are represented as thin lines for dBSE@$G_0W_0$ (blue), and dBSE@$G_0T_0$ (red) with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure*}
%%% %%% %%% %%%

@@

%%% TABLE I %%%
\begin{table*}
	\caption{Singlet and triplet excitation energies (in \si{\eV}) of various molecular systems at equilibrium geometry computed at various levels of theory with the cc-pVDZ basis set.
	The corresponding experimental equilibrium geometries have been extracted from the computational chemistry comparison and benchmark database (CCCBDB) at \url{https://cccbdb.nist.gov}.}
	\label{fig:mol}
	\begin{ruledtabular}
		\begin{tabular}{lcccccccc}
					&			&	\mc{5}{c}{Methods}																\\
			\cline{3-7}
			Mol.	&	State	&	CIS		&	CIS(D)	&	TDHF	&	BSE@$GW$	&	dBSE@$GW$	&	BSE@$GT$	&	dBSE@$GT$	\\
			\hline
			\ce{H2}	&			&	\\
			\ce{LiH}	&			&	\\
			\ce{LiF}	&			&	\\
			\ce{BF}	&			&	\\
			\ce{CO}	&			&	\\
			\ce{N2}	&			&	\\
			\ce{F2}	&			&	\\
			\ce{FH}	&			&	\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
%%% %%% %%% %%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
The authors thank Roberto Orlando for useful discussions.
PFL thanks the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~863481) for funding.
This project has also received financial support from the CNRS through the 80|Prime program and has been supported through the EUR grant NanoX ANR-17-EURE-0009 in the framework of the ``Programme des Investissements d’Avenir''.
\end{acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%
%\section{$T$-matrix self-energy in the spatial orbital basis}
%%%%%%%%%%%%%%%%%%%%%%%%

%Here, we provide implementable expressions for the $T$-matrix self-energy in the spatial orbital basis.

%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{Tmatrix}
%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In order to target neutral (singly-)excited states, we first consider the static version of the BSE employing the $GT$ quasiparticle energies [see Eq.~\eqref{eq:G0T0}] as well as $T$-matrix kernel [\ie, $\Sigma^{\GT}$ in Eq.~\eqref{eq:Xi}].
In this case, the BSE@$GT$ linear eigenvalue problem simply reads
\begin{equation}
\label{eq:BSE}
	\begin{pmatrix}
		\bA{}{\BSE}		&	\bB{}{\BSE}	\\
		-\bB{}{\BSE}	&	-\bA{}{\BSE}	\\
	\end{pmatrix}
	\cdot
	\begin{pmatrix}
		\bX{m}{\BSE}	\\
		\bY{m}{\BSE}	\\
	\end{pmatrix}
	=
	\Om{m}{\BSE}
	\begin{pmatrix}
		\bX{m}{\BSE}	\\
		\bY{m}{\BSE}	\\
	\end{pmatrix}
\end{equation}
with 
\begin{subequations}
\begin{align}
	A_{ia,jb}^{\BSE} & = \delta_{ij} \delta_{ab} (\e{a}{\GT} - \e{i}{\GT}) + \mel{ib}{}{aj} - T_{ib,aj}(\omega=0)
	\\ 
	B_{ia,jb}^{\BSE} & = \mel{ij}{}{ab} - T_{ij,ab}(\omega=0)
\end{align}
\end{subequations}
The eigenvalues $\Om{m}{\BSE}$ of Eq.~\eqref{eq:BSE} provide $OV$ singlet (\ie, spin-conserved) and $OV$ triplet  (\ie, spin-flip) single excitations.
Note that the spin structure of the BSE@$GT$ equations is analogous to the BSE@$GW$ version, \cite{Monino_2021} and one can compute separately singlet and triplet excitation energies.
Neglecting the coupling between excitations and deexcitations, \ie, $\bB{}{\BSE} = \bO$, is known as the Tamm-Dancoff approximation (TDA).

Due to the frequency-independent nature of the static BSE, it is well known that one cannot access double (and higher) excitations.
\cite{Loos_2019,Romaniello_2009b,Sangalli_2011,Loos_2020h,Authier_2020,Monino_2021}
In order to go beyond the static approximation, it is possible to consider, within the dynamical TDA (dTDA) that neglects the frequency dependence of the coupling block $\bB{}{}$, the dynamical version of the BSE (dBSE). \cite{Strinati_1988,Romaniello_2009b,Loos_2020h}
In this case, one must solve the (non-linear) dynamical eigenvalue problem
\begin{equation}
\label{eq:dBSE}
	\begin{pmatrix}
		\bA{}{\dBSE}(\Om{S}{})	&	\bB{}{\BSE}	\\
		-\bB{}{\BSE}	&	-\bA{}{\dBSE}(-\Om{S}{})	\\
	\end{pmatrix}
	\cdot
	\begin{pmatrix}
		\bX{S}{\dBSE}	\\
		\bY{S}{\dBSE}	\\
	\end{pmatrix}
	=
	\Om{S}{}
	\begin{pmatrix}
		\bX{S}{\dBSE}	\\
		\bY{S}{\dBSE}	\\
	\end{pmatrix}
\end{equation}
with 
\begin{equation}
	A_{ia,jb}^{\dBSE}(\omega) = \delta_{ij} \delta_{ab} (\e{a}{\GT} - \e{i}{\GT}) + \mel{ib}{}{aj} - \widetilde{T}_{ib,aj}(\omega)
\end{equation}
where, following Strinati's seminal work, \cite{Strinati_1988} one can derive the following expression for the elements of the dynamical $T$ matrix
\begin{equation}
\label{eq:dT}
\begin{split}
	\widetilde{T}_{pq,rs}(\omega) 
		& = \sum_{n} \frac{\braket*{pq}{\chi_n^{N+2}}\braket*{rs}{\chi_n^{N+2}}}{\omega - \Om{n}{N+2} + (\e{i}{\GT} + \e{j}{\GT}) + i\eta}
		\\
		& + \sum_{n} \frac{\braket*{pq}{\chi_n^{N-2}}\braket*{rs}{\chi_n^{N-2}}}{\omega + \Om{n}{N-2} - (\e{a}{\GT} + \e{b}{\GT}) + i\eta}
\end{split}
\end{equation}
from which, one can check that we recover the static expression \eqref{eq:T} in the limit $\Om{n}{N\pm2} \to \infty$.
Equation \eqref{eq:dT} highlights the interesting dynamical structure of the $T$ matrix, where, similarly to the dBSE@$GW$ scheme, \cite{Strinati_1988,Romaniello_2009b,Loos_2020h} the 2h2p configurations are downfolded on the 1h1p configurations. \cite{Bintrim_2021b}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%
The present formalism has been implemented in the electronic structure package \texttt{QuAcK} \cite{QuAcK} which is freely available at \url{https://github.com/pfloos/QuAcK}. 
We consider here only systems with closed-shell singlet ground states. 
Thus, the $GW$ and $GT$ calculations are performed by considering a (restricted) HF starting point and standard gaussian basis sets (defined with cartesian functions) are employed.
Note that all quasiparticle energies which are obtained via Eq.~\eqref{eq:G0T0} are corrected in the same way.
Finally, the infinitesimal $\eta$ is set to zero for all calculations.
The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here.
Although the dynamical correction is computed in the dTDA throughout, the zeroth-order excitonic Hamiltonian [see Eq.~\eqref{eq:BSE}] is always the ``full'' BSE static Hamiltonian, i.e., without TDA.

In terms of computational cost, the overall scaling of BSE@$GT$ is equivalent to BSE@$GW$ as they both correspond to seeking the lowest eigenvalues of a matrix of size $(2OV \times 2OV)$. 
Searching iteratively for the lowest eigenstates can be routinely performed via Davidson's algorithm with a $\order*{K^4}$ computational cost. \cite{Davidson_1975}
The cost of the dynamical correction, which is thoroughly discussed in Ref.~\onlinecite{Loos_2020h}, is more expensive but is again equivalent in both formalisms.
The computational cost associated with the computation of the $T$-matrix and the screening $W$ both scale as $\order*{K^6}$ as one must obtain all the eigenvalues and eigenvectors of the pp-RPA and ph-RPA problem, respectively.
However, the prefactor of the pp-RPA calculation is significantly larger than its ph-RPA counterpart due to the larger size of the pp-RPA matrices and its non-Hermitian nature. \cite{Peng_2013,Scuseria_2013,Yang_2013,Yang_2013b,Yang_2014a,Zhang_2015,Zhang_2016}
Moreover, within the $T$-matrix formalism, one must compute both the singlet and triplet contributions of the $T$-matrix, while for singlet states, only the singlet contribution of $W$ contributes to the self-energy.
Although similar approaches remain to be developed for the $T$-matrix formalism, contour deformation and density fitting techniques can be efficiently implemented in the case of $GW$ to reduce the scaling to $\order*{K^3}$. \cite{Duchemin_2019,Duchemin_2020,Duchemin_2021}

@@

%%% FIGURE 2 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{H2_gap}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure}
%%% %%% %%% %%%

@@

%%% FIGURE 3 %%%
\begin{figure*}
	\includegraphics[width=\textwidth]{BeH2}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{BeH2} as a function of the distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels.
	The dynamically-corrected BSE excitation energies are represented as thin lines for dBSE@$G_0W_0$ (blue), and dBSE@$G_0T_0$ (red) with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 4 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{BeH2_gap}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{BeH2} as a function of the internuclear distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure}
%%% %%% %%% %%%

@@

%%% TABLE I %%%
\begin{table*}
	\caption{Singlet and triplet excitation energies (in \si{\eV}) of various molecular systems at equilibrium geometry computed at various levels of theory with the cc-pVDZ basis set.
	The corresponding experimental equilibrium geometries have been extracted from the computational chemistry comparison and benchmark database (CCCBDB) at \url{https://cccbdb.nist.gov}.}
	\label{fig:mol}
	\begin{ruledtabular}
		\begin{tabular}{lcccccccc}
					&			&	\mc{7}{c}{Methods}																\\
			\cline{3-9}
			Mol.	&	State	&	CIS		&	CIS(D)	&	TDHF	&	BSE@$GW$	&	dBSE@$GW$	&	BSE@$GT$	&	dBSE@$GT$	\\
			\hline
			\ce{H2}	&			&	\\
			\ce{LiH}	&			&	\\
			\ce{LiF}	&			&	\\
			\ce{BF}	&			&	\\
			\ce{CO}	&			&	\\
			\ce{N2}	&			&	\\
			\ce{F2}	&			&	\\
			\ce{FH}	&			&	\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
%%% %%% %%% %%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
%The authors thank Roberto Orlando for useful discussions.
PFL thanks the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~863481) for funding.
This project has also received financial support from the CNRS through the 80|Prime program and has been supported through the EUR grant NanoX ANR-17-EURE-0009 in the framework of the ``Programme des Investissements d’Avenir''.
\end{acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%
\onecolumngrid
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{BSE with a dynamical kernel}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To recast the BSE with a dynamical $T$-matrix kernel as an eigenvalue equation, we follow Ref.~\onlinecite{Strinati_1980} and start from the equation for the BSE amplitude 
\begin{equation}
\label{eq:Strinati_1}
	X_S(1,1^\prime) = \int d3d4d5d6L_0(1,4,1^\prime,3)\Xi(3,5,4,6) X_S(6,5)
\end{equation}
with 
\begin{align}
\label{eq:T-matrix_kernel}
	L_0(1,4,1^\prime,3) & =-iG(1,3)G(4,1^\prime)
	&
	\Xi(3,5,4,6) & =  i\frac{\delta \Sigma(3,4)}{\delta G(6,5)}\approx -T(3,5,4,6)
\end{align}
where we consider the so-called $T$-matrix self-energy \cite{Martin_2016,Romaniello_2012} $\Sigma(3,4)=iG(7,8)T(3,8,4,7)$ with
\begin{equation}
\label{eq:T-matrix}
	T(3,8,4,7) =
	- v_c(3,8)\delta(3,4)\delta(7,8)
	+ v_c(3,8)\delta(4,8)\delta(3,7)
	+ iv_c(3,8)G(3,1')G(8,2')T(1',2',4,7)
\end{equation}  
and where we neglect the functional derivative $\delta T/\delta G$ in the kernel $\Xi$.
Making the time dependence of Eq.~\eqref{eq:Strinati_1} explicit, one gets 
\begin{multline}
\label{eqn:DYN_T}
	X_S(x_1,x_{1^\prime},\tau_{11'})e^{-i\omega_S(t_1+t_{1^\prime})/2}
	=-i\int d3d4d5d6 G(x_1,x_3;\tau_{13})G(x_4,x_{1^\prime};\tau_{11^\prime})
	\\
	\times \bar{T}(x_3,x_5,x_4,x_6;\tau_{34})X_S(x_6,x_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{multline}
where $\tau^+_{ij}=t^+_i-t_j$ with $t^+_i=t_i+\eta$ ($\eta \to 0^+$) and $T(3,5,4,6)=- \delta(\tau^+_{35})\delta(\tau^+_{64})\bar{T}(x_3,x_5,x_4,x_6;\tau_{34})$ with 
\begin{multline}
\label{eq:T_bar}
	\bar{T}(x_3,x_5,x_4,x_6;\tau_{34}) 
	= v_c(x_3,x_5)\delta(\tau_{34})\delta(x_3,x_4)\delta(x_6,x_5)-v_c(x_3,x_5)\delta(\tau_{34})	\delta(x_3,x_6)\delta(x_4,x_5)
	\\
	+ iv_c(x_3,x_5)G(x_3,x_7;\tau_{38})G(x_5,x_8;\tau^+_{38})\bar{T}(x_7,x_8,x_4,x_6;\tau_{84})
\end{multline}
Using the Fourier transform $G(\tau)=\int \frac{d\omega}{2\pi}G(\omega)e^{-i\omega\tau}$, changing variable from $t_3$ to $\tau_{34}$ and taking the limit $t_{1^\prime}=t_1^+$ we arrive at
\begin{multline}
	X_S(x_1,x_{1^\prime},0^-)
	= -i\int dx_3 dx_4 dx_5 dx_6\int d\tau_{34} \int \frac{d\nu'}{2\pi}G(x_1,x_3;\nu'+\omega_S)G(x_4,x_{1^\prime};\nu')e^{i\nu'\tau_{34}}
	\\
	 \times \bar{T}(x_3,x_5,x_4,x_6;\tau_{34})X_S(x_6,x_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}.
\end{multline}
Using the Lehman representation of the one-body Green's function in the quasiparticle approximation, i.e.
$$G(x_1,x_2;\omega)=\sum_{p}\frac{\psi_p(x_1)\psi_p^*(x_2)}{\omega-\epsilon_p+ i\eta \text{sgn}(\epsilon_p-\mu)},$$ where $\psi_p$ and $\epsilon_p$ are single-particle wavefunctions and energies, respectively, and $\mu$ the chemical potential, and multiply left- and right-hand sides by $(\epsilon_a-\epsilon_i-\omega_S)\int dx_1dx^{\prime}_1 \psi^*_a(x_1)\psi_i(x_1^\prime)$.  
Using the fact that
\begin{equation}
	\Theta(\pm\tau)e^{-i\alpha \tau}=\mp\frac{1}{2\pi i}\lim_{\eta\to 0^+}\int d\omega \frac{1}{\omega-\alpha\pm i\eta}e^{-i\omega\tau}
\end{equation}
we then arrive at
\begin{multline}
\label{eq:General_DYN}
	(\epsilon_a-\epsilon_i-\omega_S)\int dx_1dx^{\prime}_1 \psi^*_a(x_1)\psi_i(x_1^\prime)  X_S(x_1,x_{1^\prime},0^-) =
	\\
	-\int dx_3 dx_4 dx_5 dx_6\int d\tau_{34}\psi_a^*(x_3)\psi_i(x_4)
	\qty[\Theta(\tau_{34})e^{i\epsilon_{i}\tau_{34}}+\Theta(-\tau_{34})e^{i(\epsilon_{a}-\omega_S)\tau_{34}} ]
	\\
	\times \bar{T}(x_3,x_5,x_4,x_6;\tau_{34})X_S(x_6,x_5;-\tau_{34})e^{i\omega_S\tau_{34}/2}
\end{multline}

We now consider the resonant case $\omega_m>0$, for which 
\begin{equation}
\label{Eqn:chi_tau}
	X_S(x_1,x_{1^\prime},\tau_1) 
	=-e^{i\omega_S\abs{\tau_1}/2} \sum_{pq}\psi_p(x_1)\psi^*_{q}(x_{1^\prime}) \mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}
  	\qty[\Theta(\tau_1)e^{-i\epsilon_p\tau_1}+\Theta(-\tau_1)e^{-i\epsilon_{q}\tau_1} ]
\end{equation}
  where $\hat{c}^\dagger_{q}$ and $\hat{c}_p$ are creation and annihilation operators, respectively, and $\ket{N}$ and $\ket{N,S}$ are the ground state and the $S$-th excited state, respectively, of the $N$-electron system.
After some algebraic steps we arrive at 
\begin{multline}
	-(\epsilon_a-\epsilon_i-\omega_S)\mel{N}{\hat{c}^\dagger_{i}\hat{c}_a}{N,S}
	\\
    = -\sum_{pq}\mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}
    \qty{ -\frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\ \bar{T}_{iq,pa}(\omega) e^{-i\omega\eta}\left[\frac{1}{\omega_S-\omega+\epsilon_{q}+\epsilon_i+i\eta}+\frac{1}{\omega_S+\omega-\epsilon_p-\epsilon_a+i\eta}\right] }
\end{multline}
where we define 
\begin{equation}
	\bar{T}_{iq,pa}(\tau_{34})=\int dx_3x_4x_5x_6\int \psi_a^*(x_3)\psi_i(x_4)\bar{T}(x_3,x_5,x_4,x_6;\tau_{34}) \psi_p(x_6)\psi^*_{q}(x_5)
\end{equation}
Using the definition $A_S^{(qp)}=\langle N|\hat{c}^\dagger_{q}\hat{c}_p|N,S\rangle$, we arrive at

\begin{equation}
	\qty(\epsilon_a-\epsilon_i-\omega_S)A_S^{(ia)}+\sum_{pq}A_S^{(q p)}\widetilde{T}_{iq,pa}(\omega_S)=0,
\end{equation}
where we defined
\begin{equation}
	\widetilde{T}_{iq,pa}(\omega_S)
	= \qty{ \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\bar{T}_{iq,pa}(\omega)e^{-i\omega\eta} \left[\frac{1}{\omega_S-\omega+\epsilon_{q}+\epsilon_i+i\eta}+\frac{1}{\omega_S+\omega-\epsilon_p-\epsilon_a+i\eta}\right]  }
 \label{Eqn:H2preso_eigen_2_T}
\end{equation}

This is an eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as,
\begin{equation}
\label{eq:H2preso_eigen}
	H^{2p,reso}_{ia,qp}(\omega_S)A_S^{(qp)}=\omega_S A^{(ia)}_S,
\end{equation}
with
\begin{equation}
	H^{2p,reso}_{ia,qp}(\omega_S)=(\epsilon_a-\epsilon_i)\delta_{iq}\delta_{ap} + \mel{ip}{}{aq}+\widetilde{T}_{iq,pa}(\omega_S).
\end{equation}
Here $H^{2p,reso}_{ia,jb}(\omega_S)=A_{ia,jb}(\omega_S)$ and $H^{2p,reso}_{ia,bj}(\omega_S)=B_{ia,jb}(\omega_S)$.
If we drop the correlation part of $\bar{T}$ [the first two terms on the right-hand side of Eq.~\eqref{eq:T_bar}] we obtain the TDHF equations. To calculate the correlation contribution we use \alert{Eq.~(9)} in Eq.~\eqref{Eqn:H2preso_eigen_2_T}, and after integration over the frequency we arrive at \alert{Eq.~(21).}

%%%%%%%%%%%%%%%%%%%%%%%%
\twocolumngrid
\bibliography{Tmatrix}
%%%%%%%%%%%%%%%%%%%%%%%%
